---
title: "bootstrapping"
author: "Quinn Anderson"
date: "2023-11-09"
output: html_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)
library(modelr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)


theme_set(theme_minimal() + theme(legend.position = "bottom"))

set.seed(1)
```

## simulate data

```{r}
n_samp = 250

sim_df_const = 
  tibble(
    x = rnorm(n_samp, 1, 1),
    error = rnorm(n_samp, 0, 1),
    y = 2 + 3 * x + error)

sim_df_nonconst = sim_df_const |> 
  mutate(
  error = error * .75 * x,
  y = 2 + 3 * x + error)
```

Plot the datasets:

```{r}
sim_df_const |> 
  ggplot(aes(x = x, y = y)) + 
  geom_point(alpha = .5) +
  stat_smooth(method = "lm")

sim_df_nonconst |> 
  ggplot(aes(x = x, y = y)) + 
  geom_point(alpha = .5) +
  stat_smooth(method = "lm")

sim_df = 
  bind_rows(const = sim_df_const, nonconst = sim_df_nonconst, .id = "data_source")

sim_df |> 
  ggplot(aes(x = x, y = y)) + 
  geom_point(alpha = .5) +
  stat_smooth(method = "lm") +
  facet_grid(~data_source)
```

Now, let's fit simple linear regressions to both datasets. 

```{r}
lm(y ~ x, data = sim_df_const) |> broom::tidy() |> knitr::kable(digits = 3)

lm(y ~ x, data = sim_df_nonconst) |> broom::tidy() |> knitr::kable(digits = 3)
```

the way we solve this is by using the bootstrap. the usual assumptions are not met in this data set. in this case, we'll need to understand what the distribution of the intercept and slope are going to be under repeated sampling. we'll get that distribution using a bootstrap. 


## draw one bootstrap sample

```{r}
boot_sample = function(df) {
  
  sample_frac(df, replace = TRUE) |> 
    arrange(x)
  
}
```

check if this works...

```{r}
boot_sample(sim_df_nonconst) |> 
  ggplot(aes(x = x, y = y)) +
  geom_point(alpha = 0.3) +
  stat_smooth(method = "lm") +
  ylim(-5, 16)
```


## drawing many bootstrap samples

```{r}
boot_straps =
  tibble(strap_number = 1:1000,
         strap_sample = rerun(1000, boot_sample(sim_df_nonconst))
    )
```

check to see if this has worked...

```{r}
boot_straps |> 
  slice(1:3) |> 
  mutate(strap_sample = map(strap_sample, arrange, x)) |> 
  pull(strap_sample)

boot_straps |> 
  slice(1:3) |> 
  unnest(strap_sample) |> 
  ggplot(aes(x = x, y = y)) + 
  geom_point(alpha = .5) +
  stat_smooth(method = "lm", se = FALSE) +
  facet_grid(~strap_number) 
```

great! now, can I run my analysis on these...?

```{r}
boot_results = 
  boot_straps |> 
  mutate(
    models = map(.x = strap_sample, ~lm(y ~ x, data = .x)),
    results = map(models, broom::tidy)) |> 
  select(strap_number, results) |> 
  unnest(results)
  

```

what do I have now? let's look at the estimate of the sd under bootstrapping samples. what is the actual sd of the estimate under repeated sampling?

```{r}
boot_results |> 
  group_by(term) |> 
  summarize(
    mean_est = mean(estimate),
    sd_est = sd(estimate))
```

let's look at the distributions

```{r}
boot_results |>
  filter(term == "x") |> 
  ggplot(aes(x = estimate)) +
  geom_density()
```

construct bootstrap CI

```{r}
boot_results |> 
  group_by(term) |> 
  summarize(
    ci_lower = quantile(estimate, 0.025),
    ci_upper = quantile(estimate, 0.975))
```

## bootstrap using modelr

can we simplify anything?

```{r}
sim_df_nonconst |> 
  bootstrap(1000, id = "strap_number") |> 
  mutate(
    models = map(.x = strap, ~lm(y ~ x, data = .x)),
    results = map(models, broom::tidy)) |> 
  select(strap_number, results) |> 
  unnest(results) |> 
  group_by(term) |> 
  summarize(
    mean_est = mean(estimate),
    sd_est = sd(estimate))
```

check with a dataset where the assumptions are met. if your assumptions are true, you can do usual linear model stuff, but you can also do the bootstrapping procedure. 

```{r}
sim_df_const |> 
  bootstrap(1000, id = "strap_number") |> 
  mutate(
    models = map(.x = strap, ~lm(y ~ x, data = .x)),
    results = map(models, broom::tidy)) |> 
  select(strap_number, results) |> 
  unnest(results) |> 
  group_by(term) |> 
  summarize(
    mean_est = mean(estimate),
    sd_est = sd(estimate))
```














