---
title: "linear_models"
author: "Quinn Anderson"
date: "2023-11-09"
output: html_document
---

## Load and clean Airbnb data

let's set up:

```{r}
library(tidyverse)
library(p8105.datasets)

set.seed(1)
```

```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb |> 
  mutate(stars = review_scores_location / 2) |> 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) |> 
  filter(borough != "Staten Island") |> 
  select(price, stars, borough, neighborhood, room_type)
```

Now, let's fit a model !!! 

A good place to start is to consider price as an outcome that may depend on rating and borough. We fit that initial model in the following code.

```{r}
fit =
  nyc_airbnb |> 
  lm(price ~ stars + borough, data = _)
```


let's look at the fit! although these are not usually as helpful to look at as is. 

```{r}
fit
summary(fit)
summary(fit)$coef
coef(fit)
fitted.values(fit)
```

so, let's tidy the output:

```{r}
fit |> 
  broom::glance()
```


now, let's tidy up the coefficients. now it's a tibble! so we can use (manipulate) as a dataframe.

```{r}
fit |> 
  broom::tidy()
```

```{r}
fit |> 
  broom::tidy() |> 
  select(term, estimate, p.value) |> 
  mutate(term = str_replace(term, "^borough", "Borough: ")) |> 
  knitr::kable(digits = 3)
```


## Fit another model

```{r}
fit = 
  nyc_airbnb |> 
  mutate(
    borough = fct_infreq(borough),
    room_type = fct_infreq(room_type)) |> 
  lm(price ~ stars + borough + room_type, data = _)

fit |> 
  broom::tidy()
```

^ the reference categories for `borough` variable, keeping `room_type` and `stars` fixed. 



## Regression Diagnostics

the key step is to make sure that we have the residuals that we need. 

```{r}
nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  ggplot(aes(x = resid)) +
  geom_density() + 
  xlim(-100, 500)

nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  ggplot(aes(x = borough, y = resid)) +
  geom_violin()

nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  ggplot(aes(x = stars, y = resid)) +
  geom_point()
```

^ this is now a dataframe, so you can do things like plot the residuals. 

our results are also sensitive to the skewness of the p-values. we need to check that we are not violating the assumptions of non-skewed distributions, etc. 


## Hypothesis test for categorical predictor

fit a "null" and "alternative" model

```{r}
fit_null = lm(price ~ stars + borough, data = nyc_airbnb)

fit_alternative = lm(price ~ stars + borough + room_type, data = nyc_airbnb)

anova(fit_null, fit_alternative) |> 
  broom::tidy()
```

^ to test the hypothesis of whether `room_type` is significant. 



## Borough-level differences

```{r}
fit = 
  nyc_airbnb |> 
  lm(price = stars * borough + room_type * borough, data = _)

fit |> 
  broom::tidy()
```

my goal right now is to look at four different linear models, each one looking at effect of borough on stars and room type. do that using `nest`. 

also, add a linear model once you have four different dfs in one. 

```{r}
airbnb_lm = function(df) {
  lm(price ~ stars + room_type, data = df)
}

nyc_airbnb |> 
  nest(df = -borough) |> 
  mutate(
    models = map(df, airbnb_lm),
    results = map(models, broom::tidy)
  ) |> 
  select(borough, results) |> 
  unnest(results) |> 
  select(borough, term, estimate) |> 
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) |> 
  knitr::kable(digits = 2)
```


same thing but just a little different...using an anonymous function rather than creating a function as above: `\(df) lm(price ~ stars + borough, data = df)`

```{r}
nyc_airbnb |> 
  nest(df = -borough) |> 
  mutate(
    models = map(df, \(df) lm(price ~ stars + borough, data = df)),
    results = map(models, broom::tidy)
  ) |> 
  select(borough, results) |> 
  unnest(results) |> 
  select(borough, term, estimate) |> 
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) |> 
  knitr::kable(digits = 2)
```

^ don't do it if you don't have to. when you use `map` a lot, people use this more often. 


## Binary outcomes

```{r}
baltimore_df = 
  read_csv("data/homicide-data.csv") |> 
  filter(city == "Baltimore") |> 
  mutate(
    resolved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age),
    victim_race = fct_relevel(victim_race, "White")) |> 
  select(resolved, victim_age, victim_race, victim_sex)

baltimore_df |> view()
```


fitting a logistic regression


```{r}
fit_logistic = 
  baltimore_df |> 
  glm(
    resolved ~ victim_age + victim_race + victim_sex,
    data = _,
    family = binomial())
```

look at model results

```{r}
fit_logistic |> 
  broom::tidy() |> 
  mutate(OR = exp(estimate)) |> 
  select(term, estimate, OR)
```

```{r}
baltimore_df |> 
  count(victim_race) |> 
  arrange(desc(n))
```










