---
title: "linear_models"
author: "Quinn Anderson"
date: "2023-11-09"
output: html_document
---

## Load and clean Airbnb data

let's set up:

```{r}
library(tidyverse)
library(p8105.datasets)

set.seed(1)
```

```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb |> 
  mutate(stars = review_scores_location / 2) |> 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) |> 
  filter(borough != "Staten Island") |> 
  select(price, stars, borough, neighborhood, room_type)
```

Now, let's fit a model !!! 

A good place to start is to consider price as an outcome that may depend on rating and borough. We fit that initial model in the following code.

```{r}
fit =
  nyc_airbnb |> 
  lm(price ~ stars + borough, data = _)
```


let's look at the fit! although these are not usually as helpful to look at as is. 

```{r}
fit
summary(fit)
summary(fit)$coef
coef(fit)
fitted.values(fit)
```

so, let's tidy the output:

```{r}
fit |> 
  broom::glance()
```


now, let's tidy up the coefficients. now it's a tibble! so we can use (manipulate) as a dataframe.

```{r}
fit |> 
  broom::tidy()
```

```{r}
fit |> 
  broom::tidy() |> 
  select(term, estimate, p.value) |> 
  mutate(term = str_replace(term, "^borough", "Borough: ")) |> 
  knitr::kable(digits = 3)
```


## Fit another model

```{r}
fit = 
  nyc_airbnb |> 
  mutate(
    borough = fct_infreq(borough),
    room_type = fct_infreq(room_type)) |> 
  lm(price ~ stars + borough + room_type, data = _)

fit |> 
  broom::tidy()
```

^ the reference categories for `borough` variable, keeping `room_type` and `stars` fixed. 



